Paper Title*\\ {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and should not be used} \thanks{Identify applicable funding agency here. If none, delete this.}

### 2019-05-01 icdm/submission/includes/evaluation.tex
new file for evalutation and changes to tables

\begin{table}[]
\centering
\begin{tabular}{|l|l|l|l|l|}
\hline
Length of Sentences & 10 & 20 & 40 & 1 Document \\ \hline
Training Time for one batch &8min & 10min & 11min & 18min \\ \hline
Convergence Time &4 epochs & 3 epochs & 3 epochs & 3 epochs \\ \hline
Word Similarity& 0.65 & 0.66 & 0.66 & 0.66 \\ \hline
\end{tabular}
\caption{Training and Convergence time according to choice of the length of sentences in text8 dataset}
\label{table:with_20}
\end{table}

@@

where $f(w)$ is the frequency of w, and $t$ is a threshold set empirically. As Equation \ref{eq:sampling} is a probability, subsampling is not a deterministic procedure, words that may have been deleted with a threshold of $10^-2$ may stay in the dataset with a lower threshold, as shown in Table \ref{table:treshold_examples}. Mikolov et al. recommend a value between $0$ and $10^{-5}$, depending on the size of the dataset. We experimented with different values and $10^{-4}$ seemed the most suited. We did this by looking at a random set of sentences and judging the results manually. An example of the first sentence with different sampling thresholds can be found in Table \ref{table:treshold_examples}. The table shows the first 20 words of our dataset, without the words that were subsampled according to a threshold sample. Stats about subsampling can be found in Table \ref{table:treshold}.
\begin{table}[h]
\centering
\begin{tabular}{|l|l|l|l|l|l|}
\hline
Sampling Treshhold & 0 & $ 10^{-1}$&$ 10^{-2}$& $10^{-3} $ &$10^{-4} $ \\ \hline
Number of words in Dataset & 16 mio & 15mio & 11 mio & 8mio & 4 mio \\ \hline
\end{tabular}
\caption{Size of the preprocessed text8 dataset according to sampling treshold}
\label{table:treshold}
\end{table}

\begin{table}[]
\begin{tabular}{|l|l|}
\hline
Sampling Treshold & First sentence of Dataset \\ \hline
$10^{-1}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated as a term of abuse first used against early\\ working class radicals including the diggers of the english\end{tabular} \\ \hline
$10^{-2}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated as a term of abuse first used against early \\ working class radicals including diggers of english\end{tabular} \\ \hline
$10^{-3}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated a term abuse first used against early \\ working class radicals including diggers the english\end{tabular} \\ \hline
$10^{-4}$ & \begin{tabular}[c]{@{}l@{}}Anarchism originated abuse used against working class radicals\\ diggers english\end{tabular} \\ \hline
$10^{-5}$ & against radicals diggers \\ \hline
\end{tabular}
\caption{Example of a sentence with different sampling tresholds}\label{table:treshold_examples}
\end{table}

@@

\subsection{Results}
We ran multiple experiments for each optimizer. This subsection will only give an overview of the achieved results. Each subsection will give an explanation over the achieved result with a specific optimizer.

\subsubsection{SGD}
The first challenge for each optimizer was to find a correct learning rate. As SGD is the optimizer used in Gensim \cite{gensim} we first tried the same learning rate as Gensim \cite{gensim} and then performed a random search to find a better one. As expected a bell curve shape resulted, a learning rate that is too high leads to diversion and a learning rate that is too low leads to a training time that is too slow. The best value that we found for the learning rate is $0.0075$. With this setting SGD converged in 11 epochs. The second experiment was to add input shuffling.
As seen in Figure \ref{fig:results_sgd}, for almost every learning rate the convergence time decreased. Our model, with the best setting, now converges in only 7 epochs. Another interesting fact to point out from Figure \ref{fig:results_sgd} is that with input shuffling we achieved better results with higher learning rates. As for learning rates of $0.01$ and $0.025$ we did converge in 11 epochs with input shuffling but did not converge in 20 epochs without it.

\begin{figure}[h]
\centering
\includegraphics[scale=0.45]{images/results_sgd_shuffle}
\caption{Training time Stochastic Gradient Descent with input Shuffling}
\label{fig:results_sgd}
\end{figure}
\subsubsection{Momentum and Nesterov}
Momentum and NAG \cite{nag} both have an additional hyperparameter $\gamma$, that, as described in Section \ref{optimizers}, defines the percentage of the previous gradient that will be added to the current gradients. We set $\gamma = 0.9$ as this is the typical value and did not alter it during our experiments. Momentum and Nesterov alone respectively only slightly decrease or increase the convergence time. Momentum optimally converges in 9 epochs and Nesterov in 13. If we combine these optimizers with input shuffling, interestingly the same phenomena as with plain SGD appear. The convergence time gets better, 8 epochs for Momentum and 3 epochs for NAG. The phenomena that a higher learning rat yields better results also happens with both of the optimizers. As Momentum does not converge in 20 epochs with a learning rate of 0.002 but does in 8 with input shuffling.
\begin{figure}[h]
\centering
\begin{minipage}{.5\textwidth}
\centering
\includegraphics[scale=0.3]{images/results_mom_shuffle}
\caption{Training time Momentum with input Shuffling}
\label{fig:results_mom}
\label{fig:test1}
\end{minipage}%
\begin{minipage}{.5\textwidth}
\centering
\includegraphics[scale=0.3]{images/results_nag_shuffle}
\caption{Training time Nesterov with input Shuffling}
\label{fig:results_nag}
\end{minipage}
\end{figure}
\subsubsection{Adagrad}
Adagrad \cite{adagrad} is a very interesting tool for learning word embeddings as it decreases the learning rate for very frequent occurring features, and vice versa (this is explained in detail in Section \ref{ssec:adagrad}). Because words that appear very frequently often do not have a semantic gain that is as important as words that appear less frequently to their context words, it's good to have a lower learning rate. So, in theory, Adagrad is particularly well suited for our task. This was confirmed empirically as our model converged in 4 epochs. When combined with shuffling Adagrad only took 3 epochs to converge. This shows the tendency of the skip gram model to converge faster with input shuffling and the big impact of having different learning rate for each feature.
Here it's interesting to notice that a higher learning rate combined with input shuffling did not yield better results than without shuffling. Both of our best results happened with a learning rate of $0.1$, as shown in Figure \ref{fig:results_adagrad_shuffle}.
\begin{figure}[h]
\centering
\includegraphics[scale=0.45]{images/results_adagrad_shuffle}
\caption{Training time Adagrad with input Shuffling}
\label{fig:results_adagrad_shuffle}
\end{figure}
\subsubsection{Adadelta}
In theory Adadelta \cite{adadelta} should outperform Adagrad as it's an extension of the former. Because it didn't have any learning rate to tune, we only did 2 experiments, with and without input shuffling. We are aware of the fact that there are additional hyperparameters but did, for simplicity reason, and because their effect is not as high as the learning rate in other optimizers, decide not to tune it. We left it to its default value $\rho = 0.9$. This parameter defines the percentage taken when calculating the exponentially decaying average of past gradients, as explained in \ref{ssec:adadelta} Adadelta did not manage to achieve a word similarity of 0.66. It only converged to a similarity of 0.59. It did this in 20 epochs without input shuffling and in 3 with input shuffling, as can be seen in Table \ref{table:results_adadelta}
\begin{table}[h]
\centering
\begin{tabular}{|l|l|l|}
\hline

\end{tabular}
\caption{}
\label{}
\end{table}

\begin{table}[tb]
    \caption{Convergence Time and Quality with Adadelta}
    %\scriptsize
    \begin{tabular}{l r r }%
        \toprule
Adadelta Model & Convergence Time & Word similarity \\ 
        \midrule%
        Without Shuffling & 20 & 0.59 \\ 
With Shuffling & 3 & 0.59 \\
        \midrule%
   \end{tabular}%
   \label{table:results_adadelta}%
\end{table}

\subsubsection{Adam}
Adam is the most advanced of all the optimizers used in our experiments and did yield the best results as seen in Figure \ref{fig:results_adam_shuffle}. Adam converged in 3 epochs without shuffling and 2 with. This is the best result that we got with any optimizer.  It's also interesting to note that as same as with Adagrad it did not react to input shuffling the same way as SGD did. In fact, it worked in the opposite direction, as we achieved our best result with input shuffling while having a lower learning rate $0.001$ then we used to achieve the best result without input shuffling $0.05$.
\begin{figure}[h]
    \centering
            \includegraphics[scale=0.45]{images/results_adam_shuffle} 
    \caption{Training time Adam with input Shuffling}
    \label{fig:results_adam_shuffle}
\end{figure}

>>>>

\begin{table}[tb]
    \caption{Training and Convergence time according to choice of the length of sentences in text8 datasetCaption}
    %\scriptsize
    \begin{tabular}{l r r r r}%
        \toprule
          Length of Sentences & 10 & 20 & 40 & 1 Document \\

        \midrule%
Training Time for one batch &8min & 10min & 11min & 18min \\ 
Convergence Time in epochs &4  & 3  & 3  & 3  \\ 
Word Similarity& 0.65 & 0.66 & 0.66 & 0.66 \\ \hline
        \midrule%
   \end{tabular}%
   \label{table:with_20}%
\end{table}

@@

where $f(w)$ is the frequency of w, and $t$ is a threshold set empirically. As Equation \ref{eq:sampling} is a probability, subsampling is not a deterministic procedure, words that may have been deleted with a threshold of $10^-2$ may stay in the dataset with a lower threshold, as shown in Table \ref{table:treshold_examples}. Mikolov et al. recommend a value between $0$ and $10^{-5}$, depending on the size of the dataset. We experimented with different values and $10^{-4}$ seemed the most suited. We did this by looking at a random set of sentences and judging the results manually. An example of the first sentence with different sampling thresholds can be found in Table \ref{table:treshold_examples}. The table shows the first 20 words of our dataset, without the words that were subsampled according to a threshold sample. Stats about subsampling can be found in Table \ref{table:treshold}.
\begin{table}[h]
\centering
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\end{tabular}

\label{table:treshold}
\end{table}

\begin{table}[tb]
\caption{Size of the preprocessed text8 dataset according to sampling treshold}
    %\scriptsize
    \begin{tabular}{l r r r r r}%
        \toprule
Sampling Treshhold & 0 & $ 10^{-1}$&$ 10^{-2}$& $10^{-3} $ &$10^{-4} $ \\

        \midrule%
Words in Dataset  & 16mio & 15mio & 11mio & 8mio & 4mio \\ 
        \midrule%
   \end{tabular}%
   \label{table:with_20}%
\end{table}

\begin{table}[tb]
\caption{Example of a sentence with different sampling tresholds}
    %\scriptsize
    \begin{tabular}{l l}%
        \toprule
Sampling Treshhold & 0 \\

        \midrule%
Sampling Treshold & First sentence of Dataset \\ \hline
$10^{-1}$ &Anarchism originated as a term of abuse first\tabularnewline used against early working class radicals \tabularnewline including the diggers of the english \\ \hline
$10^{-2}$ & Anarchism originated as a term of abuse first used against early \tabularnewline working class radicals including diggers of english \\ \hline
$10^{-3}$ & Anarchism originated a term abuse first used against early \tabularnewline working class radicals including diggers the english\\ \hline
$10^{-4}$ & Anarchism originated abuse used against working class radicals\tabularnewline diggers english \\ \hline
$10^{-5}$ & against radicals diggers \\ \hline        
        \midrule%
   \end{tabular}%
\label{table:treshold_examples}
\end{table}

@@
